---
title: "EDA - Looking at the data"
output: html_notebook
---

Pasting my script so that it can be displayed with the output

```{r, echo=TRUE}
# Exploration and initial ideas:

# Load data ----

library(readxl)
library(dplyr)

scores_un <- read_xlsx("Data/panel_case_data_DS_II_2023.xlsx", sheet = "SCORES BY CHILD - unpivot")
scores_piv <- read_xlsx("Data/panel_case_data_DS_II_2023.xlsx", sheet = "SCORES BY CHILD")

regions <- read_xlsx("Data/panel_case_data_DS_II_2023.xlsx", sheet = "REGIONS & BASES")
var_names_tbl <- read_xlsx("Data/panel_case_data_DS_II_2023.xlsx", sheet = "VARIABLE NAMES")


## Renaming for ease: ----

# This is a bit more intuitive for me
scores_un_renamed <- scores_un %>% rename(
    math_adv_score = EDU_ICAN_ADVANCED,
    math_basic = EDU_ICAN_FOUNDNUM,
    reading_drop = EDU_MICS_LIT_ENDED,
    reading_understanding = EDU_MICS_LIT_FOUND,
    participated_tvet = EDU_NF_TVET_PART,
    is_enrolled_school = EDU_PIE_IN_SCHOOL,
    received_certification = EDU_PIE_TVET_CERT,
    is_enrolled_tvet = EDU_PIE_TVET_PART,
    thinking_score = EDU_SOFT_HOTS_AVG,
    communication_score = EDU_SOFT_SCS_AVG,
    is_working = EDU_WORK_EMPLOY_YES,
    not_s_w_t = EDU_WORK_NEET,
    is_working_field = EDU_WORK_TVETCONN,
    )
````
# EDA ----

# Profiling with the DataExplorer package

```{r, eval=FALSE}
library(DataExplorer)

# No need to run again. Outputs are in the Output folder
scores_piv %>%
    create_report(
        output_file = "profile_report_unpivoted_data",
        output_dir = "Outputs/",
        y = "SCORE",
        report_title = "EDA for Students Self-Sufficiency scores - pivoted data"
    )

scores_un_renamed %>%
    create_report(
        output_file = "profile_report_pivoted_data",
        output_dir = "Outputs/",
        y = "is_working",
        report_title = "EDA for Students Self-Sufficiency scores - unpivoted data"
    )
```




### Initial reactions(thinking out loud): ----

1) Missing data
a) Reason?
Are certain regions not collecting certain data ? Why is there so much missing data?
Or it may be according to age group ? May take a look and also talk
I think I should break by age group, as the variables seem to be more relevant per age group (e.g. working)
Comments after investigating: Yes, the main reason is because certain variables are only relevant for certain age groups.
b) Dealing with it
Also, there might be values that may be imputed, such as obs where they are working in the field and so, they are also working. (is_working_field, is_working)
Such as
There are certain variables that are related to each other.
2) Breaking down:
Visualizing by region may be useful

# Disaggregate by age group for profiling ----

```{r}
scores_un_renamed %>% select(AGE_GROUP) %>% unique() # Getting the age groups

```


```{r, echo=FALSE}
# Ages 19+
scores_un_renamed_19up <- scores_un_renamed %>% 
    filter(AGE_GROUP == "19+") 
# Ages 15-18
scores_un_renamed_15_18 <- scores_un_renamed %>% 
    filter(AGE_GROUP == "15-18") 
# Ages 12-14
scores_un_renamed_12_14 <- scores_un_renamed %>% 
    filter(AGE_GROUP == "12-14") 
# Ages 9-11
scores_un_renamed_9_11 <- scores_un_renamed %>% 
    filter(AGE_GROUP == "9-11") 
# Ages 6-8
scores_un_renamed_6_8 <- scores_un_renamed %>% 
    filter(AGE_GROUP == "6-8") 
```



# SKIMR ----

```{r}
library(skimr)
skim(scores_un_renamed_19up) 
```

Ok, so I found out that reading_understanding, math_basic, reading_drop and math_adv_score are not captured for this age group. Which makes total sense. 

```{r}
skim(scores_un_renamed_15_18)
```

Lower rates of completion are the participated_tvet, is_working_field (which is dep in tvet* check), and received_certification 
 
```{r}
skim(scores_un_renamed_12_14) 
```
 
Same as 15_18

```{r}
skim(scores_un_renamed_9_11)
```

No data for not_s_w_t , participated_tvet, is_enrolled_tvet, is_working, is_working_field, received_certification, communication_score, thinking_score

```{r}
skim(scores_un_renamed_6_8)
```
Solely has data for is_enrolled_school. And 99.7% of them are.

# Jotting down some insights/thoughts/ideas:
Can compare reading_understanding with EGRA scores for East Africa ?


Ok, I think it is time to wrap up this EDA file and start the outputs in the analysis.
There are more things I'd love to explore, such as if using factor analysis may be relevant, as well as some clustering and statistical tests. But I think I better first get a good descriptive picture with the interpretable ones.
I also think it is relevant to include data from East Africa to contextualize the findings. If I have time, I will.

